### 2017-04-24 notes/initial_writeup/paper.tex
no message

\documentclass[sigconf]{acmart}

% BEGIN ACM HEADER

\usepackage{booktabs} % For formal tables
\usepackage{balance} % For balanced columns on the last page

% Copyright
\setcopyright{acmlicensed}

% DOI - Replace your DOI below...
\acmDOI{}

% ISBN - Replace your conference/workshop's ISBN below...
\acmISBN{}

% SELECT YOUR CONFERENCE/WORKSHOP BELOW...
\acmConference[]{}{}{}

\acmYear{2017}
\copyrightyear{2017}

\acmPrice{}

% END ACM HEADER

\usepackage{epstopdf}
\usepackage{url}

\usepackage{tikz}
\usetikzlibrary{shadows,shapes}
\usepackage{pgfplots}
\pgfplotsset{compat=1.10}

\usepackage{xcolor}
\usepackage{colortbl}
\definecolor{Gray}{gray}{0.90}

\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

\usepackage[textsize=small]{todonotes}
\newcommand{\TODO}[1]{\todo[inline]{#1}}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\vect}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\vpi}[0]{\vect{\pi}}

\newcommand{\calA}[0]{\ensuremath{\mathcal{A}}}
\newcommand{\calC}[0]{\ensuremath{\mathcal{C}}}
\newcommand{\calS}[0]{\ensuremath{\mathcal{S}}}

\begin{document}

\title[Blockchain as Middleware Services for Transactive Energy Applications]{Blockchain as Middleware Services for\\Transactive Energy Applications}

\author{Aron Laszka}
\affiliation{%
  \institution{Vanderbilt University}
  \city{Nashville} 
  \state{Tennessee} 
  \postcode{37212}
}
\email{aron.laszka@vanderbilt.edu}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{A. Laszka et al.}

\begin{abstract}
We develop a set of powerful and novel distributed market middleware services that enable local prosumers (a consumer with renewable generation capability) to trade with each other, as well as to bid for power resources, while following the suggested timing constraints.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002951.10003227</concept_id>
<concept_desc>Information systems~Information systems applications</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003462.10003544</concept_id>
<concept_desc>Social and professional topics~Commerce policy</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10010583.10010662.10010663</concept_id>
<concept_desc>Hardware~Energy generation and storage</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10010583.10010662.10010674</concept_id>
<concept_desc>Hardware~Power estimation and optimization</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{Information systems~Information systems applications}
\ccsdesc[300]{Social and professional topics~Commerce policy}
\ccsdesc[300]{Hardware~Energy generation and storage}
\ccsdesc[300]{Hardware~Power estimation and optimization}

\keywords{Transactive microgrid, blockchain}

\maketitle

%\section{Introduction}

\section{Architecture Overview}

\begin{figure}[h!]
\begin{tikzpicture}[
  x=1.04cm,
  nodeStyle/.style={rounded corners=0.1cm, drop shadow={shadow xshift=0.04cm, shadow yshift=-0.05cm, fill=black}}]

\draw [fill=blue!5, nodeStyle] (0, 3.07) rectangle (1.55, 3.93) node [midway, align=center, font=\small] {Trading\\[-0.2em]Node};
\draw [fill=blue!5, nodeStyle] (1.65, 3.07) rectangle (3.15, 3.93) node [midway, align=center, font=\small] {Ledger\\[-0.2em]Node};
\draw [fill=blue!5, nodeStyle] (3.25, 3.07) rectangle (4.75, 3.93) node [midway, align=center, font=\small] {Prosumer\\[-0.2em]Node};
\draw [fill=blue!5, nodeStyle] (4.85, 3.07) rectangle (6.35, 3.93) node [midway, align=center, font=\small] {Smart Meter\\[-0.2em]Node};
\draw [fill=blue!5, nodeStyle] (6.45, 3.07) rectangle (8.0, 3.93) node [midway, align=center, font=\small] {DSO\\[-0.2em]Node};

\draw [fill=red!5, nodeStyle] (0, 2.07) rectangle (3.95, 2.93) node [midway] {Trading Platform};
\draw [fill=red!5, nodeStyle] (4.05, 2.07) rectangle (8, 2.93) node [midway] {Distributed Ledger};

\draw [fill=green!5, nodeStyle] (0, 1.07) rectangle (3.95, 1.93) node [midway] {RIAPS};
\draw [fill=green!5, nodeStyle] (4.05, 1.07) rectangle (8, 1.93) node [midway] {Cryptographic Libraries};

\draw [fill=black!5, nodeStyle] (0, 0.07) rectangle (8, 0.93) node [midway] {Operating System};
\end{tikzpicture}
\caption{High-level software architecture.}
\label{fig:softwareArchitecture}
\end{figure}

Figure~\ref{fig:softwareArchitecture} shows a high-level overview of the software architecture of our system.
Our system is -- in part -- built on the Resilient Information Architecture for Smart Grid (RIAPS) platform~\cite{eisele2017riaps}.

Figure~\ref{fig:logicalNodeArchitecture} shows a high-level overview of the trading platform, the distributed ledger, and the logical nodes implementing and interacting with these subsystems.

\paragraph{Trading Platform}
Although individual prosumers trade energy with each other, for the sake of scalability, we need a trading platform that enables prosumers to find trade partners.
By serving as a database for energy bids, the trading platform relieves prosumer nodes from communicating with all potential trade partners.~%
\footnote{Note that prosumers are not required to use the services of the trading platform, and they are allowed to communicate with each other for the sake of finding trade partners.
However, in a microgrids with a large number prosumers, the trading platform will be able to find partners more efficiently and promptly.}
In addition to receiving and storing bids from prosumers, the trading platform can also find matches and notify the prosumers of the opportunity to make an energy trade.
Even though the trading platform appears to the nodes as one entity, for the sake of scalability and reliability, it can be implemented in a distributed manner, using multiple trading nodes.

\paragraph{Distributed Ledger}
The distributed ledger stores transactions that specify energy trades, record energy production and consumption, and change regulatory policies for the microgrid, including energy prices.
Since the ledger is maintained by multiple ledger nodes, a key implementation requirement is reaching consensus on which transactions are valid and stored in the ledger.
This consensus must be reached quickly, reliably, and securely even in the presence of erroneous or even malicious ledger nodes.
We treat the consensus algorithm as a pluggable module, which can be based on blockchains with proof-of-stake consensus or Practical Byzantine Fault Tolerance algorithm~\cite{castro1999practical}.

\begin{figure}[h!]
\begin{tikzpicture}[
  x=1.6cm,
  font=\small,
  nodeStyle/.style={fill=white, draw=black, drop shadow=black!30, minimum size=0.5cm},
  arrowStyle/.style={->, >=stealth, shorten >=1.5pt, shorten <=1.5pt},
  transactionStyle/.style={draw, align=center, fill=black!10, font=\footnotesize}]
\draw [->, >=stealth] (0.6, 0) -- (5.4, 0) node [right] {time};

\foreach \pos in {1, 2, 3, 4, 5}
  \draw (\pos, 0.06cm) -- (\pos, -0.06cm);

\foreach \pos/\name in {{1/$T-2$}, {2/$T-1$}, {3/$T$}, {4/$T+1$}}
  \draw[decorate, decoration={brace,amplitude=0.15cm}, yshift=-0.12cm] ({\pos+0.95}, 0) -- ({\pos+0.05}, 0) node [midway, below=0.15cm, font=\small] {\name};

\node [nodeStyle, rounded corners=0.1cm] (prosumer) at (1.5, 1.5) {};
\node [nodeStyle, diamond] (smartmeter) at (4.5, 1.5) {};

\node [above=-0.05cm of prosumer] {prosumer};
\node [above=0cm of smartmeter] {smartmeter};

\node [transactionStyle] (financial) at (1.5, 0.5) {financial\\[-0.25em]transaction};
\node [transactionStyle] (energy) at (4.5, 0.5) {energy\\[-0.25em]transaction};

\draw [arrowStyle] (prosumer) -- (financial);
\draw [arrowStyle] (smartmeter) -- (energy);

\draw [arrowStyle, dashed] (financial) -- (3.25, 0.5) node [midway, above] {promise} -- (3.25, 0);
\draw [arrowStyle, dashed] (energy) -- (3.5, 0.5) node [midway, above] {record} -- (3.5, 0);
\end{tikzpicture}
\caption{Basic timeline of transactions.}
\label{fig:timeline}
\end{figure}

Time is divided into fixed-length intervals, and energy trades are specified with respect to these intervals: financial transactions promise to consume or produce a certain amount of energy in a future interval, while energy transactions record the actual amount of energy produced or consumed in an earlier interval. 
Figure~\ref{fig:timeline} shows a basic timeline of financial and energy transactions.

\begin{figure*}
\begin{tikzpicture}[
  x=1cm, 
  y=1cm,
  nodeStyle/.style={fill=white, draw=black, drop shadow=black!30, minimum size=0.5cm}]

% trading nodes
\begin{scope}[xshift=1cm]

\foreach \name/\x/\y in {{t1/0/0}, {t2/1/1.4}, {t3/2/0}}
  \node [regular polygon, regular polygon sides=5, nodeStyle] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{tAnchorRight/2.5/0.7}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{t1/t2}, {t2/t3}, {t3/t1}}
  \draw [<->, >=stealth, shorten >=1.5pt, shorten <=1.5pt] (\from) -- (\to);

\draw [decorate, decoration={brace,amplitude=0.2cm}, yshift=-0.4cm] (2.25, 0) -- (-0.25, 0) node [midway, below=0.2cm] {trading nodes};
\end{scope}

% prosumer nodes
\begin{scope}[xshift=6cm]

\foreach \name/\x/\y in {{p1/0/0}, {p2/1.4/0}, {p3/0/1.4}, {p4/1.4/1.4}}
  \node [nodeStyle, rounded corners=0.1cm] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{pAnchorLeft/-0.6/0.7}, {pAnchorRight/2/0.7}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{p1/p2}, {p1/p3}, {p1/p4}, {p2/p3}, {p2/p4}, {p3/p4}}
  \draw [<->, >=stealth, shorten >=1.5pt, shorten <=1.5pt, dashed] (\from) -- (\to);

\draw [decorate, decoration={brace,amplitude=0.2cm}, yshift=-0.4cm] (1.65, 0) -- (-0.25, 0) node [midway, below=0.2cm] {prosumer nodes};

\end{scope}

% smart-meter nodes
\begin{scope}[xshift=6cm, yshift=2.7cm]

\foreach \name/\x/\y in {{m1/0/0}, {m2/1.4/0}, {m3/0/1.4}, {m4/1.4/1.4}}
  \node [diamond, nodeStyle] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{mAnchorRight/2/0.7}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{m1/m2}, {m1/m3}, {m1/m4}, {m2/m3}, {m2/m4}, {m3/m4}}
  \draw [<->, >=stealth, shorten >=1.5pt, shorten <=1.5pt, dashed] (\from) -- (\to);

\draw [decorate, decoration={brace,amplitude=0.2cm}, yshift=0.4cm] (-0.25, 1.4) -- (1.65, 1.4) node [midway, above=0.2cm] {smart-meter nodes};

\end{scope}

% ledger nodes
\begin{scope}[xshift=11.5cm]

\foreach \name/\x/\y in {{l1/0/0}, {l2/1/1.4}, {l3/2/0}}
  \node [circle, nodeStyle] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{lAnchorLeft/-0.5/0.7}, {lAnchorTopLeft/0/1.6}, {lAnchorTopRight/1.6/1.0}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{l1/l2}, {l2/l3}, {l3/l1}}
  \draw [<->, >=stealth, shorten >=1.5pt, shorten <=1.5pt] (\from) -- (\to);

\draw [decorate, decoration={brace,amplitude=0.2cm}, yshift=-0.4cm] (2.25, 0) -- (-0.25, 0) node [midway, below=0.2cm] {ledger nodes};

\end{scope}

% additional nodes

\node [ellipse, nodeStyle] (dso) at (15.5cm, 3.0cm) {DSO node};

\draw [->, very thick, >=stealth, bend right=10, shorten <=0.5cm, shorten >=0.5cm] (dso) to node [midway, above left=0.0cm, align=center] {regulatory\\[-0.25em]transactions} (lAnchorTopRight);
\draw [<-, very thick, >=stealth, bend left=10, shorten <=0.5cm, shorten >=0.5cm] (dso) to node [midway, below right=0.0cm, align=center] {ledger status} (lAnchorTopRight);

% connections between systems
\draw [<-, very thick, >=stealth, bend left=5] ([yshift=0.0cm]tAnchorRight) to node [midway, above=0.1cm] {bids} ([yshift=0.0cm]pAnchorLeft);
\draw [->, very thick, >=stealth, bend right=5] ([yshift=-0.4cm]tAnchorRight) to node [midway, below=0.1cm] {matches} ([yshift=-0.4cm]pAnchorLeft);

\draw [->, very thick, >=stealth, bend left=5] ([yshift=0.0cm]pAnchorRight) to node [midway, above=0.1cm, align=center] {financial\\[-0.25em]transactions} ([yshift=0.0cm]lAnchorLeft);
\draw [<-, very thick, >=stealth, bend right=5] ([yshift=-0.4cm]pAnchorRight) to node [midway, below=0.1cm] {ledger status} ([yshift=-0.4cm]lAnchorLeft);

\draw [->, very thick, >=stealth, bend left=15] ([yshift=0.2cm]mAnchorRight) to node [midway, above right=0.0cm, align=center] {energy\\[-0.25em]transactions} ([yshift=0.2cm]lAnchorTopLeft);
\draw [<-, very thick, >=stealth, bend right=-5] ([yshift=-0.2cm]mAnchorRight) to node [midway, below left=0.0cm] {ledger status} ([yshift=-0.2cm]lAnchorTopLeft);

\end{tikzpicture}
\caption{High-level logical architecture of the trading platform and the distributed ledger. A physical node may implement multiple logical nodes.}
\label{fig:logicalNodeArchitecture}
\end{figure*}

\begin{table*}
\caption{Logical Node Types}
\label{tab:logicalNodeTypes}
\begin{tabular}{|c||l|l|l|}
\hline
\textbf{Type} & \textbf{Function} & \textbf{Cardinality} & \textbf{Physical node} \\
\hline\hline
\rowcolor{Gray} Ledger      & maintain distributed ledger                 & one or more           & any node (with enough computational power) \\ \hline
                Trading     & provide trading platform                    & one or more           & any node (with enough computational power) \\ \hline
\rowcolor{Gray} Prosumer    & trade electricity (on behalf of a prosumer) & one for each prosumer & smart inverter or smart meter \\ \hline
                Smart meter & record power consumption and production     & one for each prosumer & smart meter (must be trusted hardware) \\ \hline
\rowcolor{Gray} DSO         & regulate microgrid electricity market       & one                   & node owned by the DSO \\ \hline
\end{tabular}
\end{table*}

\subsection{Logical Nodes}

Next, we discuss the five types of logical nodes in our system: ledger, trading, prosumer, smart meter, and DSO nodes.
Table~\ref{tab:logicalNodeTypes} shows an overview of these logical-node types.

Every node has a pair of public and private keys as well as a digital certificate from the DSO, which proves the ownership of the public key.
The nodes can use the key pair and the certificate prove their identity and to sign messages or transactions.
A digital certificate may also specify the capabilities of a node.
For example, the certificate of a prosumer node may specify the amount of energy that the node is allowed to trade for a single time interval.

\subsubsection{Ledger nodes}
Ledger nodes implement the distributed ledger by forming a peer-to-peer network with each other and reaching a consensus on the state of the ledger.
A ledger node is responsible for receiving financial transactions from prosumer nodes, energy transactions from smart-meter nodes, and regulatory transactions from the DSO node.
The ledger nodes share these transactions with each other, verify them, and reach a consensus on which transactions are recorded in the ledger.

\subsubsection{Trading nodes}
Trading nodes implement the distributed trading platform by forming a peer-to-peer network with each other and providing services for prosumer nodes.
A trading node is responsible for receiving bids from prosumer nodes, sharing these bids with the other trading nodes, maintaining a distributed database of active bids, trying to match bids with each other, and notifying the prosumers of matches.
While this platform is similar to the distributed ledger in many ways, it is subject to much less stringent requirements.
For example, the trading platform does not have to store history in a reliable and verifiable manner.
Consequently, trading nodes can be implemented by computationally less capable physical nodes (e.g., with limited storage).

\begin{table*}
\caption{Transaction Types}
\label{tab:transactionTypes}
\begin{tabular}{|c||l|l|p{5.8cm}|}
\hline
\textbf{Type} & \textbf{Creator} & \textbf{Contents} & \textbf{Validity} \\
\hline\hline
\rowcolor{Gray} Financial  & prosumers    & identifiers, signatures, time interval, amount of energy & signed by buyer and seller, abides by the safety and security rules \\ \hline
                Energy     & smart meters & identifier, signature, time interval, amount of energy   & signed by smart meter \\ \hline
\rowcolor{Gray} Regulatory & DSO          & signature, time interval, set of policy updates            & signed by DSO \\ \hline
\end{tabular}
\end{table*}

\subsubsection{Prosumer nodes}
Each prosumer node represents a prosumer, such as a residential customer.
A prosumer node trades future electricity consumption and production on behalf of the prosumer.
As a first step in electricity trading, the prosumer node estimates the net energy demand of the prosumer for a future time interval.
At the same time, it queries the active regulatory policies, such as energy prices, from the distributed ledger.
Based on the estimated net demand and the active regulatory policies, it creates an energy trading bid, which it then submits to the trading platform.
If the trading platform responds with one (or more) matching bid, then the prosumer node contacts the node that submitted a matching bid.
The two nodes negotiate, agree on, create, and sign a new financial transaction, which they then submit to the distributed ledger.
Once the ledger has verified and recorded the transaction, the electricity trade has been set up.

\subsubsection{Smart-meter nodes}
Each smart-meter node measures the actual energy production and consumption of a prosumer.
For each time interval, the smart-meter node measures the net energy consumption of the prosumer, records the net consumption in an energy transaction, and then signs the transaction and submits it to the distributed ledger.
To ensure that the smart-meter node measures the correct amount of net consumption, the node must always be implemented by a trusted smart meter deployed at the prosumer.
Furthermore, we assume that the smart-meter devices are tamper-proof hardware, which prevents tampering attacks by the prosumers (e.g., to avoid paying for electricity).

\subsubsection{DSO node}
The DSO node represents the Distribution System Operator, and it is responsible for regulating the microgrid market using regulatory transactions.
These transactions enable the DSO, for example, to set a price policy for the microgrid or to remove misbehaving nodes by revoking their certificates.
Note that in order to enhance reliability and security, the DSO node may be split into multiple logical nodes.
For example, the DSO node can be split into a pricing node, which is authorized and responsible for setting the price policy, and a revocation node, which is authorized and responsible for revoking digital certificates.

\subsection{Transactions}

Transactions on the distributed ledger can be divided into three main types: financial, energy, and regulatory transactions.
Table~\ref{tab:transactionTypes} shows an overview of these transaction types.

\subsubsection{Financial Transactions}
A financial transaction represents an agreement between two prosumers, called the seller and the buyer, who promise to respectively produce and consume a certain amount of energy in a future time interval.

A financial transaction contains the following fields:
\begin{itemize}
\item \textbf{Seller ID}: Unique identifier of the selling prosumer node.
\item \textbf{Buyer ID}: Unique identifier of the buying prosumer node.
\item \textbf{Monetary value}: Value paid by the buyer to the seller, which can be denominated in either a fiat currency (e.g., US dollars) or a cryptocurrency.
\item \textbf{Time interval}: Identifier of the time interval in which energy will be produced and consumed.
\item \textbf{Energy value}: Amount of energy produced by the seller and consumed by the buyer.
\item \textbf{Seller signature}: Digital signature created using the private key of the seller, computed over all previous fields.
\item \textbf{Buyer signature}: Digital signature created using the private key of the seller, computed over all previous fields except the seller signature.
\end{itemize}

The ledger nodes accept and record a financial transaction if
\begin{itemize}
\item both signatures are valid,
\item time interval is in the future,
\item transaction abides by the safety and security rules set by the DSO (we will discuss these rules in detail later).
\end{itemize}

\subsubsection{Energy Transactions}
An energy transaction records the actual amount of energy consumed and produced by a prosumer in a past time interval.

An energy transaction contains the following fields:
\begin{itemize}
\item \textbf{Smart-meter ID}: Unique identifier of the smart meter.
\item \textbf{Energy value}: Net amount of energy consumed by the prosumer, which is computed as the difference between the consumption and the production of the prosumer (note that it may be a negative value if the prosumer produced more than its consumption).
\item \textbf{Time interval}: Identifier of the time interval in which energy consumption and production were measured.
\item \textbf{Smart-meter signature}: Digital signature created using the private key of the smart meter, computed over all previous fields.
\end{itemize}

The ledger nodes accept and record an energy transaction if
\begin{itemize}
\item signature is valid
\item time interval is in the past.
\end{itemize}

\subsubsection{Regulatory Transactions}
A regulatory transaction changes the set of active policies which all transactions are subject to.
To provide elaborate tools for regulating a microgrid, our system may support a variety of intricate regulatory transactions.
Here, we discuss two transactions that are necessary for regulating any microgrid market: price-policy and certificate-revocation transactions.

\paragraph{Price policy}
A price-policy transaction changes the active price policy of the DSO for the microgrid.
A price policy specifies how much a prosumer will have to pay to the DSO for the difference between its promised and actual net consumption.

A price-policy transaction contains the following fields:
\begin{itemize}
\item \textbf{Time interval}: Identifier of the time interval from which the new price policy will be active.
\item \textbf{Price policy}: Specification of the new price policy.
\item \textbf{DSO signature}: Digital signature created using the private key of the DSO, computed over all previous fields.
\end{itemize}

The ledger nodes accept and record a price-policy transaction if
\begin{itemize}
\item signature is valid
\item time interval is in the future.
\end{itemize}

\paragraph{Certificate Revocation}
A certificate-revocation transaction revokes some of the digital certificates that were issued to the nodes.
After a certificate has been revoked, ledger nodes will no longer accept digital signatures created using key pairs whose ownership was proven by the certificate.

A certificate-revocation transaction contains the following fields:
\begin{itemize}
\item \textbf{Time interval}: Identifier of the time interval in which the certificate is revoked.
\item \textbf{List of certificates}: Certificates to be revoked.
\item \textbf{DSO signature}: Digital signature created using the private key of the DSO, computed over all previous fields.
\end{itemize}

The ledger nodes accept and record a certificate-revocation transaction if
\begin{itemize}
\item signature is valid
\item time interval is in the future.
\end{itemize}

\section{Related Work}

Microsoft Azure Blockchain as a Service (BaaS) and Project Bletchley~\cite{gray2016introducing}.

Hyperledger Fabric~\cite{hyperledger2017fabric}.

Interledger~\cite{thomas_protocol}.

Bitcoin Lightning Network~\cite{poon2016bitcoin}.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}