### 2017-05-15 papers/IoT2017/figArchitecture.tex
IoT 2017 paper started

\begin{figure*}
\begin{tikzpicture}[
  x=1cm, 
  y=1cm,
  nodeStyle/.style={fill=white, draw=black, drop shadow=black!30, minimum size=0.5cm},
  linkStyle/.style={<->, >=stealth, shorten >=1.5pt, shorten <=1.5pt},
  batchStyle/.style={->, very thick, >=stealth},
  braceStyle/.style={decorate, decoration={brace,amplitude=0.2cm}, yshift=-0.4cm},
]

% trading nodes
\begin{scope}[xshift=1cm]

\foreach \name/\x/\y in {{t1/0/0}, {t2/1/1.4}, {t3/2/0}}
  \node [nodeStyle, regular polygon, regular polygon sides=5] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{tRight/2.5/0.7}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{t1/t2}, {t2/t3}, {t3/t1}}
  \draw [linkStyle] (\from) -- (\to);

\draw [braceStyle] (2.25, 0) -- (-0.25, 0) node [midway, below=0.2cm] {trading nodes};
\end{scope}

% prosumer nodes
\begin{scope}[xshift=6cm]

\foreach \name/\x/\y in {{p1/0/0}, {p2/1.4/0}, {p3/0/1.4}, {p4/1.4/1.4}}
  \node [nodeStyle, rounded corners=0.1cm] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{pLeft/-0.6/0.7}, {pRight/2/0.7}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{p1/p2}, {p1/p3}, {p1/p4}, {p2/p3}, {p2/p4}, {p3/p4}}
  \draw [linkStyle, dashed] (\from) -- (\to);

\draw [braceStyle] (1.65, 0) -- (-0.25, 0) node [midway, below=0.2cm] {prosumer nodes};
\end{scope}

% smart-meter nodes
\begin{scope}[xshift=6cm, yshift=2.7cm]

\foreach \name/\x/\y in {{m1/0/0}, {m2/1.4/0}, {m3/0/1.4}, {m4/1.4/1.4}}
  \node [diamond, nodeStyle] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{mRight/2/0.7}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{m1/m2}, {m1/m3}, {m1/m4}, {m2/m3}, {m2/m4}, {m3/m4}}
  \draw [linkStyle, dashed] (\from) -- (\to);

\draw [braceStyle, yshift=0.8cm] (-0.25, 1.4) -- (1.65, 1.4) node [midway, above=0.2cm] {smart-meter nodes};
\end{scope}

% ledger nodes
\begin{scope}[xshift=11.5cm]

\foreach \name/\x/\y in {{l1/0/0}, {l2/1/1.4}, {l3/2/0}}
  \node [circle, nodeStyle] (\name) at (\x,\y) {};

\foreach \name/\x/\y in {{lLeft/-0.5/0.7}, {lTopLeft/0/1.6}, {lTopRight/1.6/1.0}}
  \coordinate (\name) at (\x,\y);

\foreach \from/\to in {{l1/l2}, {l2/l3}, {l3/l1}}
  \draw [<->, >=stealth, shorten >=1.5pt, shorten <=1.5pt] (\from) -- (\to);

\draw [decorate, decoration={brace,amplitude=0.2cm}, yshift=-0.4cm] (2.25, 0) -- (-0.25, 0) node [midway, below=0.2cm] {ledger nodes};
\end{scope}

% additional nodes
\node [ellipse, nodeStyle] (dso) at (15cm, 3.1cm) {DSO node};

\draw [->, >=stealth, bend right=10, shorten <=0.25cm, shorten >=0.5cm] (dso) to node [midway, above left=0cm, align=center] {regulatory\\[-0.25em]transactions} (lTopRight);
\draw [<-, >=stealth, bend left=10, shorten <=0.25cm, shorten >=0.5cm] (dso) to node [midway, below right=0cm, align=center] {ledger status} (lTopRight);

\node [above=1.2cm of dso] (controller) {control algorithm};

\draw [<-, >=stealth, bend right=10, shorten <=1.5pt] (dso) to node [midway, left=0.3cm, align=center] {measured and\\[-0.25em]expected load} (controller);
\draw [->, >=stealth, bend left=10, shorten <=1.5pt] (dso) to node [midway, right=0.3cm, align=center] {control\\[-0.25em]signal} (controller);

% connections between systems
\draw [batchStyle, <-, bend left=5] (tRight) to node [midway, above=0.1cm] {bids} (pLeft);
\draw [batchStyle, bend right=5] ([yshift=-0.4cm]tRight) to node [midway, below=0.1cm] {matches} ([yshift=-0.4cm]pLeft);

\draw [batchStyle, bend left=5] (pRight) to node [midway, above=0.1cm, align=center] {financial\\[-0.25em]transactions} ([yshift=0cm]lLeft);
\draw [batchStyle, <-, bend right=5] ([yshift=-0.4cm]pRight) to node [midway, below=0.1cm] {ledger status} ([yshift=-0.4cm]lLeft);

\draw [batchStyle, bend left=15] ([yshift=0.2cm]mRight) to node [midway, above right=0cm, align=center] {energy\\[-0.25em]transactions} ([yshift=0.2cm]lTopLeft);
\draw [batchStyle, <-, bend right=-5] ([yshift=-0.2cm]mRight) to node [midway, below left=0cm] {ledger status} ([yshift=-0.2cm]lTopLeft);

\end{tikzpicture}
\caption{High-level logical architecture of the trading platform and the distributed ledger. A physical node may implement multiple logical nodes.}
\label{fig:logicalNodeArchitecture}
\end{figure*}