### 2017-06-05 papers/IoT2017/paper.tex
no message

\documentclass[chi_draft]{sigchi}

% Use this section to set the ACM copyright statement (e.g. for
% preprints).  Consult the conference website for the camera-ready
% copyright statement.

% Copyright
\CopyrightYear{2017}
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}
% DOI
\doi{} %http://dx.doi.org/10.475/123_4}
% ISBN
\isbn{} %123-4567-24-567/08/06}
%Conference
\conferenceinfo{IoT'17,}{October 22--25, 2017, Linz, Austria}
%Price
\acmPrice{\$15.00}

% Use this command to override the default ACM copyright statement
% (e.g. for preprints).  Consult the conference website for the
% camera-ready copyright statement.

%% HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP --
%% Please note you need to make sure the copy for your specific
%% license is used here!
% \toappear{
% Permission to make digital or hard copies of all or part of this work
% for personal or classroom use is granted without fee provided that
% copies are not made or distributed for profit or commercial advantage
% and that copies bear this notice and the full citation on the first
% page. Copyrights for components of this work owned by others than ACM
% must be honored. Abstracting with credit is permitted. To copy
% otherwise, or republish, to post on servers or to redistribute to
% lists, requires prior specific permission and/or a fee. Request
% permissions from \href{mailto:Permissions@acm.org}{Permissions@acm.org}. \\
% \emph{CHI '16},  May 07--12, 2016, San Jose, CA, USA \\
% ACM xxx-x-xxxx-xxxx-x/xx/xx\ldots \$15.00 \\
% DOI: \url{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
% }

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy
% \pagenumbering{arabic}

% Load basic packages
\usepackage{balance}       % to better equalize the last page
\usepackage{graphics}      % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}

% Some optional stuff you might like/need.
\usepackage{microtype}        % Improved Tracking and Kerning
% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
\usepackage{ccicons}          % Cite your images correctly!
% \usepackage[utf8]{inputenc} % for a UTF8 editor only

% Custom packages

\usepackage{paralist}
\usepackage{enumitem}

\usepackage{tikz}
\usetikzlibrary{shadows,shapes}
\usepackage{pgfplots}
\pgfplotsset{compat=1.10}

\usepackage{xcolor}
\usepackage{colortbl}
\definecolor{Gray}{gray}{0.90}

\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

% If you want to use todo notes, marginpars etc. during creation of
% your draft document, you have to enable the "chi_draft" option for
% the document class. To do this, change the very first line to:
% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
% by using the "\todo{...}"  command. Make sure to disable the draft
% option again before submitting your final document.
\usepackage[textsize=tiny, textwidth=1.3cm]{todonotes}
\newcommand{\TODO}[1]{\todo[inline]{#1}}
\newcommand{\Abhishek}[1]{\todo[color=yellow!50, linecolor=black!50]{\textbf{Abhishek}: #1}}
\newcommand{\Aron}[1]{\todo[color=green!40, linecolor=black!50]{\textbf{Aron}: #1}}

\newcommand{\circled}[1]{\tikz[baseline=(char.base)]{\node[shape=circle, draw, inner sep=0.5pt, solid] (char) {#1};}}

\newcommand{\field}[1]{\texttt{#1}}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Private Blockchains for Transactive Energy:\\Security, Privacy, and Safety}
\def\plainauthor{First Author, Second Author, Third Author}
\def\emptyauthor{}
\def\plainkeywords{Internet of Things; blockchain; transactive energy; privacy; security; transactive microgrid; smart grid; anonymity.}
\def\plaingeneralterms{Design, Human Factors, Reliability, Security} % TODO: revise

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  pdfdisplaydoctitle=true, % For Accessibility
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\allowdisplaybreaks

\title{\plaintitle}

\numberofauthors{3}
\author{%
%  \alignauthor{Leave Authors Anonymous\\
%    \affaddr{for Submission}\\
%    \affaddr{City, Country}\\
%    \email{e-mail address}}\\
%  \alignauthor{Leave Authors Anonymous\\
%    \affaddr{for Submission}\\
%    \affaddr{City, Country}\\
%    \email{e-mail address}}\\
%  \alignauthor{Leave Authors Anonymous\\
%    \affaddr{for Submission}\\
%    \affaddr{City, Country}\\
%    \email{e-mail address}}\\
}

\maketitle

\begin{abstract}
Power grids are undergoing major changes due to rapid growth in renewable energy resources and improvements in battery technology.
While these changes enhance sustainability and efficiency, they also create significant management challenges as the complexity of power systems increases.
To tackle these challenges, decentralized IoT solutions are emerging, which may arrange local communities into transactive microgrids.
Within a transactive microgrid, consumers with energy generation and storage capabilities may trade energy with each other, thereby smoothing the load on the main grid using local supply.
However, providing security, safety, and privacy in a decentralized and transactive energy system is very challenging.
On the one hand, consumers' personal information must be protected from their trade partners and \Abhishek{To what extent is this different that the information already known by the system operator? Is this more related to the ``real-time'' nature of transactive information? Rather than aggregative total information which is available at the end of billing cycle today.} \Aron{Do we need to detail this in the abstract?} the system operator.
On the other hand, the system must be protected from careless or malicious trading, which could destabilize the entire grid. 
In this paper, we design a secure and safe solution for transactive microgrids, which enables consumers to trade energy without sacrificing their privacy.
Our solution builds on distributed ledgers, such as blockchains, and provides anonymity for communication, bidding, and trading.
\end{abstract}

% TODO!
\category{K.6.m}{Miscellaneous}{Security}
\category{D.4.7}{Organization and Design}{Distributed systems}

\keywords{\plainkeywords}

\input{intro.tex}

\input{prelim.tex}

\input{solution.tex}

\section{Discussion}
\label{sec:discussion}

Here, we present a semi-formal analysis of the proposed solution, and show that it satisfies the security, safety, and privacy requirements.

\subsection{Security}

Most of our security requirements follow readily from 
\begin{compactitem}
\item the non-malleability of transactions in the distributed ledger 
\item the signature based validation of transactions
\item tamper-resistance of smart meters.
\end{compactitem}

\subsection{Safety}

First, we can easily see that the net amount of energy sold by a prosumer for each timestep is at most $\field{MAXEPA}_i$.
The amount of energy sold by the prosumer for a timetep is less than or equal to the amount of EPA obtained for that timestep.
The prosumer can withdraw $\field{MAXEPA}_i$ amount of EPA from its smart meter.
However, to obtain any amount of further EPA, the prosumer has to buy energy from another prosumer, which would decrease its net energy sold by the same amount.
Similarly, the net amount of energy sold is at most $\field{MAXECA}_i$.

Second, we can see that the amount of bids posted at the same time by prosumer $i$ for timestep $t$ is at most $\field{MAXEPA}_i + \field{MAXECA}_i$.

\subsection{Privacy}

Due to the communication anonymity and mixing services, members of a microgrid can see only the amount of energy and financial assets withdrawn by a prosumer.
Since all trading transactions are anonymous, the amount of assets traded by the prosumer will not be publicly known.
In case a prosumer does not wish to trade, it can anonymously deposit its assets to a random address that was freshly generated by its smart meter.
Note that even if a prosumer does not wish to trade, it should always withdraw and mix assets; otherwise, the lack of withdrawal would leak information.

As for the DSO, it can receive the same information from the smart meter as in a non-transactive grid (i.e., amount of energy produced and consumed).
\Aron{Also knows financial amounts, but those are necessary for billing.}
Since trading is anonymous, it does not receive any further information.

In fact, we can provide an even higher-level of privacy.
Since price policies are recorded on the ledger, which the smart meters may read, the smart meters can calculate the prosumer's monthly bill, and send only this single financial amount to the DSO each month.
Meanwhile, the DSO can still obtain current and future load information from the bid storage and from the trades recorded on the ledger.

\input{related.tex}

\input{concl.tex}

% BALANCE COLUMNS
\balance{}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}

\end{document}