### 2017-06-06 papers/IoT2017/related.tex
Update of Related Work + 1 citation addition

\section{Related Work}
\label{sec:related}

\subsection{Smart Grid and Meter Privacy}

% http://ieeexplore.ieee.org/abstract/document/5054916/
%McDaniel and McLaughlin discuss privacy challenges in smart grids~\cite{mcdaniel2009security}.

% https://arxiv.org/pdf/1108.2234.pdf
% https://pdfs.semanticscholar.org/cdf8/a5b6256823bca38a1d2347ab36f8e4a2ca94.pdf

% http://www.comm.toronto.edu/~akhisti/sm.pdf

% https://www.researchgate.net/profile/Georgios_Kalogridis/publication/224189766_Smart_Grid_Privacy_via_Anonymization_of_Smart_Metering_Data/links/541169510cf2b4da1bec4193.pdf

% https://arxiv.org/pdf/1305.0735.pdf

Both old and new security and privacy concerns arise with the continuing adoption of smart grids. In addition to both old and new security threats, such as energy theft and smart meter malware, McDaniel and McLaughlin discuss the privacy concerns of energy usage profiling that smart grids could potentially enable\cite{mcdaniel2009security}.

Several approaches have been investigated as potential means to provide privacy protections for energy consumers. Some approaches look to the use of protocols and/or frameworks to help protect privacy. Rajagopalan et al.\ use tools from information theory to present a framework, which abstracts both the privacy and the utility requirements of smart meter data~\cite{rajagopalan2011smart,sankar2013smart}. Their framework leads to a novel tractable privacy-utility tradeoff problem with minimal assumptions. Efthymiou and Kalogridis describe a method for securely anonymizing frequent electrical metering data sent by a smart meter~\cite{efthymiou2010smart}. 
Their approach is based on the observation that frequent metering data may be required by an energy distribution network for operational reasons, but it may not necessarily need to be attributable to a specific smart meter.
The authors describe a method that provides a third-party escrow mechanism for authenticated anonymous meter readings, which are difficult to associate with particular smart meter.

Additionally, other approaches such as additional hardware components are explored for potential privacy gains. Varodayan and Khisti study using rechargeable battery for partially protecting the privacy of information contained in a household's electrical load profile~\cite{varodayan2011smart}.
They show that stochastic battery policies may leak 26\% less information than a best-effort policy, holds the output load constant whenever possible. Tan et al. study privacy in a smart metering system from an information theoretic perspective in the presence of energy harvesting and storage units\cite{tan2013increasing}. 
They show that energy harvesting provides increased privacy by diversifying the energy source, while a storage device can be used to increase both energy efficiency and privacy. 
They show that there exists a trade-off between the information leakage rate and the wasted energy rate, and study the impact of the energy harvesting rate and the size of the storage device on this trade-off.

\subsection{Blockchain Technology}

As Blockchain technology continues to develop and mature, new frameworks, services, and protocols are being developed to leverage blockchain's distributed ledger. Microsoft offers Blockchain as a Service (BaaS) on Azure. Additionally, Microsoft has Project Bletchley as its architectural approach to building an Enterprise Consortium Blockchain Ecosystem, introducing two new concepts: blockchain middleware and cryptlets a secure means for calling code or data outside a SmartContract or blockchain~\cite{gray2016introducing}. Hyperledger Fabric is a platform for distributed ledger solutions, which was designed to support pluggable implementations of different components~\cite{hyperledger2017fabric}. Bitcoin Lightning Network is decentralized system, in which transactions are sent over a network of micropayment channels whose transfer of value occurs off-blockchain~\cite{poon2016bitcoin}. Interledger is a protocol for payments across payment systems, which enables anyone with accounts on two ledgers to create
a connection between them~\cite{thomas_protocol}. Xu et al.\ discuss the use of two pools of proxy agents, agreement pool and payment pool, to assist in protection of privacy when using blockchain technologies for transactions on tangible goods.  \cite{Xu2017}

%\url{https://geli.net/residential/}
%\url{https://www.greentechmedia.com/articles/read/geli-raises-7m-to-take-energy-storage-software-to-the-next-level}

%\url{http://ethembedded.com/}