### 2017-06-16 papers/IoT2017/services.tex

\subsection{Services}

We now describe the various services that are provided in our system.
%on which transactions are built and which build on transactions.
Earlier, we discussed the distributed ledger, which permanently stores
valid transactions.  Below, we introduce the anonymous communication
service, the mixing service for transaction anonymity, the anonymous
bid storage, and smart-meter based billing.

\subsubsection{Communication Anonymity}
The anonymous communication layer is the infrastructure upon which all
other services in our system are built.  Without this communication
layer, transactions and bids could be easily de-anonymized based on
their sources' network identifiers (\emph{e.g.}, IP or MAC addresses).

We can employ well-known and widely used techniques for anonymous
communication, such as \emph{onion routing}~\cite{reed1998anonymous}.
To build an onion network, the smart meters, prosumers, and other
devices can act as onion routers, and the list of onion routers in a
microgrid can be published on the ledger.  In practice, this service 
can be built on 
%In our first implementation, we can use
the free and open-source Tor software with private Directory
Authorities.  In this case, anonymous communication identifiers in
bids and asks correspond to public-keys that identify Tor hidden
services.
% ritter.vg: "run your own tor network"
% https://ritter.vg/blog-run_your_own_tor_network.html

\subsubsection{Transaction Anonymity}
Communication anonymity is necessary, but not sufficient, for
anonymous trading. In particular, if prosumers used their own accounts
to transfer assets, their trades would not be anonymous.  Fortunately,
most distributed ledgers allow users to easily generate new addresses
%\footnote{The usage of the term \emph{address} varies between distributed ledgers, but our system could be implemented using any popular ledger, such as Bitcoin and Ethereum.
%Specifically, we use the term address to denote a possible destination for asset transfers. 
%Assets transferred to an address can be used only by someone who knows the private key of the address (typically, the one who generated the address).} 
at random, which are anonymous in the sense that no one can tell who
generated them.  However, if prosumers simply transferred assets to these
addresses, they could be easily de-anonymized by tracing the
assets back to the prosumers.

To prevent this de-anonymization, prosumers transfer assets to their
anonymous addresses through a \emph{mixing service}.  The mixing
service prevents tracing assets back to their original owners by
mixing together multiple incoming transfers and multiple outgoing
transfers. This service thus hides the connections between the
prosumers and the anonymous addresses.

A mixing service can be implemented using multiple approaches.  The
simplest one is to use a \emph{trusted third party}, called a
cryptocurrency tumbler, which can receive and send assets. However, anonymity
in this case depends on the trustworthiness and reliability
of the third party, who could easily de-anonymize the addresses.  A
more secure approach is to use decentralized protocols, such as
CoinShuffle~\cite{ruffing2014coinshuffle} or
Xim~\cite{bissias2014sybil}.  These protocols enable participants to
mix assets with each other, thereby eliminating the need for a trusted
third party.
%, which would constitute a single point of failure.
Some newer cryptocurrencies, such as
Zerocoin~\cite{miers2013zerocoin}, provide built-in mixing services,
which are often based on cryptographic principles and proofs.

\begin{comment}
We must provide prosumers with the ability to create and publish
transactions anonymously.  More specifically, prosumers should be able
to purchase or sell energy without revealing their identity; however,
these transaction must also be verifiable and enforceable.

We may achieve this goal using multiple approaches for blockchain transaction anonymity:
\begin{itemize}
\item Mixing services (also known as tumblers) mix potentially
  identifiable assets on a blockchain with others, thereby preventing
  tracing individual assets back to their original source.  In our
  case, assets to be mixed include virtual balances of fiat currencies
  as well as energy production and consumption.
\item Cryptographic anonymity for transactions is provided, for
  example, by Zerocoin~\cite{miers2013zerocoin}. Similarly to mixing
  service, Zerocoin can prevent tracing assets on a blockchain.
\end{itemize}
Using the above techniques, we can enable prosumers to trade energy
anonymously (i.e., without revealing their true identities), but at
the same time prevent them from altering their energy or financial
balances without a valid transaction.

However, we must also ensure that 1) smart meters know the amount of
energy purchased or sold by their prosumer and 2) prosumers cannot
purchase or sell more energy than their capacity.  To satisfy both of
these constraints, all energy trades must start with the prosumer
withdrawing a certain amount of energy production or consumption from
its smart meter:
\begin{itemize}
\item If a prosumer wishes to sell energy, it must first withdraw an
  energy asset from its smart meter using a blockchain transaction.
  This transaction must be signed by the smart meter, which enables
  the smart meter to 1) keep track of the amount of energy traded by
  the prosumer as well as to 2) enforce safety requirements by
  limiting the amount of energy that can be withdrawn.
\item If a prosumer wishes to buy energy, it must first withdraw an
  energy consumption asset from its smart meter in a way similar to
  withdrawing an energy production asset.
\end{itemize}
\end{comment}

\subsubsection{Bidding Anonymity}
Prosumers must be able to anonymously post energy bids and asks on the
bid storage service.  An anonymous bid (or ask) contains an ECA (or
EPA), a price, and an anonymous communication identifier (\emph{e.g.},
Tor hidden service), which can be used to contact the bidding (or
asking) prosumer.  To enforce safety requirements, the bid storage
service must verify that the prosumer actually owns the asset to be
traded.  To this end, the prosumer first has to prove that it controls
the anonymous address where the asset is stored, which can be
performed in multiple ways.

In many distributed ledgers, an address represents a public key, and
controlling means knowing the corresponding private key.  In this
case, the prosumer can prove that it controls an address by signing a
challenge, which was freshly generated by the service, with the
private key of the address.  Alternatively, the prosumer may also
prove control by transferring zero amount of assets to a random
address that was freshly generated by the service.

%Finally, we must provide prosumers with the ability to post energy bids and asks anonymously.
%To this end, we create a storage for anonymous bids that is readable by all the prosumers in the microgrid.
%Any prosumer may submit a bid to this storage; however, in order to do so, they must provide a zero-knowledge proof of owning the assets that are to be traded:
%\begin{itemize}
%\item To submit an energy sell bid, the prosumer must prove that it owns an energy production asset on the chain.
%\item To submit an energy buy bid, the prosumer must prove that it own an energy consumption asset as well as financial assets on the chain.
%\end{itemize}

\subsubsection{Smart-Meter Based Billing}
After a prosumer has finished trading, it deposits all of its EPA,
ECA, and FA to the smart meter by transferring them to an anonymous
address generated by the smart meter.  Later, during timeslot $t$, the
smart meter measures the amount of energy actually consumed (or
produced) by the prosumer using physical sensors.  The meter can then
compute the prosumer's bill for timeslot $t$, which will be paid to
the DSO, as follows.

The energy consumption balance $E_i^t$ of prosumer $i$ is
\begin{align*}
E_i^t = & \hphantom{+} \text{measured net energy consumption during timeslot } t \\ %\text{measured consumption} - \text{measured production} \\
 & - \sum_{epa \,\in\, \{\text{EPA deposited by } i\}: ~ epa.\field{start} \,\leq\, t \,\leq\, epa.\field{end}} epa.\field{power} \\
 & + \sum_{epa \,\in\, \{\text{EPA withdrawn by } i\}: ~ epa.\field{start} \,\leq\, t \,\leq\, epa.\field{end}} epa.\field{power} .
% & + \text{EPA deposited by $i$} - \text{EPA withdrawn by $i$} .
\end{align*}
Notice that consumption assets are not used directly for billing, they are only used to enforce security and safety requirements.

The bill $B_i^t$ of prosumer $i$ for timeslot $t$, which will be paid
by the prosumer to the DSO, is
\begin{align*}
B_i^t = &  \text{FA withdrawn by $i$ during $t$} - \text{FA deposited by $i$ during $t$} \\
 & + \begin{cases}
- E_i^t \cdot \field{priceProduction} & \text{ if } E_i^t < 0 \\
 E_i^t \cdot \field{priceConsumption} & \text{ otherwise,} 
\end{cases}
\end{align*}
where {\field{priceProduction}} and {\field{priceConsumption}} are the
prices set by the latest regulatory transactions for timeslot $t$.