### 2017-06-16 papers/IoT2017/solution.tex

\section{Proposed System}

This section describes our solution for providing privacy to prosumers
in a transactive energy IoT system, without compromising grid safety
and security.
\Doug{Abhishek, could this figure appear earlier, i.e., in the previous section, as per my comment three?}
Figure~\ref{fig:softwareArchitecture} shows the architecture of 
\Doug{Abhishek, can we give ``our solution'' a name so that we can refer to it by name throughout the paper?}
our solution.  In this figure, components marked red are introduced to
provide privacy in a safe and secure manner, whereas components marked
in blue are typical elements of a decentralized transactive microgrid.

\begin{figure}[h!]
\center
\begin{tikzpicture}[x=8cm, y=1.1cm, font=\small,
  nodeStyle/.style={rounded corners=0.1cm, drop shadow={shadow xshift=0.05cm, shadow yshift=-0.05cm, fill=black}}
]
%\draw [nodeStyle, fill=white]   (0, 0) rectangle    (1, 0.9) node [midway, align=center] {Communication network\\[-0.2em]and operating system};

%\draw [nodeStyle, fill=red!10]  (0, 1) rectangle    (1, 1.9) node [midway, align=center] {Communication anonymity\\[-0.2em](e.g., onion routing)};

%\draw [nodeStyle, fill=blue!10] (0, 2) rectangle    (1, 2.9) node [midway, align=center] {Distributed ledger\\[-0.2em](e.g., blockchain)};

%\draw [nodeStyle, fill=red!10]  (0,    3) rectangle (0.45, 3.9) node [midway, align=center] {Transaction anonymity\\[-0.2em](mixing service)};
%\draw [nodeStyle, fill=blue!10] (0.47, 3) rectangle (0.72, 3.9) node [midway, align=center] {Bid storage};
%\draw [nodeStyle, fill=red!10 ] (0.74, 3) rectangle (1,    3.9) node [midway, align=center] {Active\\[-0.2em]smart meter};

%\draw [nodeStyle, fill=red!10] (0,    4) rectangle (0.49, 4.9) node [midway, align=center] {Prosumer anonymous\\[-0.2em]trading process};
%\draw [nodeStyle, fill=blue!10] (0.51, 4) rectangle (1,    4.9) node [midway, align=center] {Microgrid\\[-0.2em]controller};
\draw [nodeStyle, fill=white]   (0, 0.9) rectangle    (1, 1.5) node [midway, align=center] {Communication network and operating system};

\draw [nodeStyle, fill=red!10]  (0, 1.65) rectangle    (1, 2.25) node [midway, align=center] {Communication anonymity (e.g., onion routing)};

\draw [nodeStyle, fill=blue!10] (0, 2.4) rectangle    (1, 3.0) node [midway, align=center] {Distributed ledger (e.g., blockchain)};

\draw [nodeStyle, fill=red!10]  (0, 3.15) rectangle (0.45, 4.05) node [midway, align=center] {Transaction anonymity\\[-0.2em](mixing service)};
\draw [nodeStyle, fill=blue!10] (0.47, 3.15) rectangle (0.72, 4.05) node [midway, align=center] {Bid storage};
\draw [nodeStyle, fill=red!10 ] (0.74, 3.15) rectangle (1, 4.05) node [midway, align=center] {Active\\[-0.2em]smart meter};

\draw [nodeStyle, fill=red!10] (0, 4.2) rectangle (0.49, 5.1) node [midway, align=center] {Anonymous trading\\[-0.2em]process for prosumers};
\draw [nodeStyle, fill=blue!10] (0.51, 4.2) rectangle (1, 5.1) node [midway, align=center] {Microgrid controller};
\end{tikzpicture}
\caption{High-level architecture of our solution.}
\label{fig:softwareArchitecture}
\end{figure}

\begin{figure*}[h]
\center
\begin{tikzpicture}[x=3.8cm, y=1.6cm, font=\small,
  system/.style={draw, align=center, rounded corners=0.1cm, fill=black!5},
  entity/.style={draw, align=center, rounded corners=0.1cm},
  asset/.style={midway, align=center},
  transfer/.style={->, >=stealth, shorten <=0.05cm, shorten >=0.05cm},
]
%\node[entity] (smartmeter) at (0.75, 0.5) {smart\\meter};
%\node[entity] (prosumer1) at (1.25, 1) {prosumer};
%\node[system] (mixing1) at (2, 1) {mixing\\service};
%\node[entity] (prosumer2) at (3, 1) {prosumer\\(alternative address)};
%\node[system] (bidstorage) at (4, 1) {bid\\storage};
%\node[entity] (partner) at (4, 0) {other prosumer};
%\node[entity] (prosumer3) at (3, 0) {prosumer\\(alternative address)};
%\node[system] (mixing2) at (2, 0) {mixing\\service};
%\node[entity] (prosumer4) at (1.25, 0) {prosumer};
\node[entity] (smartmeter) at (0, 1) {smart meter};
\node[entity] (prosumer1) at (1, 1) {prosumer};
\node[system] (mixing1) at (1.9, 1) {mixing\\service};
\node[entity] (prosumer2) at (3, 1) {prosumer\\(anonymous address)};
\node[system] (bidstorage) at (4, 0.5) {bid\\storage};
\node[entity] (partner) at (3, 0) {other prosumer};
\node[entity] (prosumer3) at (1.5, 0) {prosumer\\(anonymous address)};
\node[entity] (smartmeter2) at (0, 0) {smart meter\\(anonymous address)};

%\draw[transfer] (smartmeter) -- node [asset, above left] {energy\\production asset} (prosumer1);
%\draw[transfer] (prosumer1) -- node [asset, above] {energy\\production asset} (mixing1);
%\draw[transfer] (mixing1) -- node [asset, above] {energy\\production asset} (prosumer2);
%\draw[transfer, dashed] (prosumer2) -- node [asset, above] {energy production asset,\\energy bid} (bidstorage);
%\draw[transfer, dashed] (bidstorage) -- node [asset, right] {energy\\bid} (partner);
%\draw[transfer, bend right=15] (partner) to node [asset, below] {financial asset,\\energy\\consumption asset} (prosumer3);
%\draw[transfer, bend right=7.5] (prosumer2) to node [asset, above right] {energy\\prod. asset} (partner);
%\draw[transfer] (prosumer3) -- node [asset, below] {financial asset,\\
%energy consumption asset} (mixing2);
%\draw[transfer] (mixing2) -- node [asset, below] {financial asset,\\energy consumption asset} (prosumer4);
%\draw[transfer] (prosumer4) -- node [asset, below left] {financial\\asset} (smartmeter);

\draw[transfer] (smartmeter) -- node [asset, above] {\circled{1} energy\\production asset} (prosumer1);
\draw[transfer] (prosumer1) -- node [asset, above] {\circled{2} energy\\production asset} (mixing1);
\draw[transfer] (mixing1) -- node [asset, above] {\circled{3} energy\\production asset} (prosumer2);
\draw[transfer, dashed] (prosumer2) -- node [asset, above right, xshift=-1.5em, yshift=0.5em] {\circled{4} energy production asset,\\energy ask} (bidstorage);
\draw[transfer, dashed] (bidstorage) -- node [asset, below right] {energy\\ask} (partner);
\draw[transfer, bend right=30] (partner) to node [asset, below] {\circled{5} financial asset,\\energy consumption asset} (prosumer3);
\draw[transfer, bend right=50] (prosumer2) to node [asset, right, yshift=0.25em] {\circled{5} energy\\production asset} (partner);
\draw[transfer] (prosumer3) -- node [asset, below] {\circled{6} financial asset,\\
energy consumption asset} (smartmeter2);
\end{tikzpicture}
\caption{Simplified overview of the flow of assets from the perspective of a prosumer who sells energy.
Note that to prevent de-anonymization, a prosumer should use multiple addresses and multiple rounds of mixing, which we have omitted from the figure for ease of presentation.}
\label{fig:sellFlow}
\end{figure*}

\subsection{Overview of the Trading Process}
We now provide a semi-formal description of the energy trading process
from the prosumers' perspective. 
Subsequent subsections will describe the assets, transactions, and services used for trading in more detail.

First, consider a prosumer who wishes to sell energy to another
prosumer, as shown in Figure~\ref{fig:sellFlow}. 
As its first step, the prosumer withdraws an \emph{energy production asset} from its smart meter.
An energy production asset represents a permission to sell a certain amount of energy, and it is used to enforce safety requirements.
If the prosumer has sufficient unsold production capacity, the smart meter creates and transfers a production asset to the prosumer using a \emph{smart meter transaction} \circled{1}, which is recorded on the distributed ledger.

At this point, the production asset can still be traced back to the prosumer since the ledger is public.
To achieve anonymity, the prosumer uses a \emph{mixing service}, which could be implemented as a decentralized protocol, such as CoinShuffle~\cite{ruffing2014coinshuffle} or Xim~\cite{bissias2014sybil}.
In turn, the mixing service
transfers the production asset to an \emph{anonymous address} \circled{3},
%\Aron{We should probably insert a good definition here for readers who are unfamiliar with blockchain transactions.}
which is randomly generated and controlled by the
prosumer.\footnote{The meaning of \emph{address} varies between distributed ledgers, but our system could be implemented using any popular blockchain, such as Bitcoin and Ethereum. 
Specifically, we use the term address to denote a possible destination for asset transfers. 
Assets that have been transferred to an address can be used only by someone who ``controls'' the address (typically, the one who generated it), which usually means knowing a private key that corresponds to the address.} % knows the private key of the address (typically, the one who generated the address).}
% knows the private key of the address (typically, the one who generated the address).}
Since the mixing service transfers assets from multiple prosumers to
multiple anonymous addresses at the same time, and the anonymous
addresses were generated at random by the prosumers, the assets
cannot be traced back to the original prosumers after
mixing.\footnote{Note that prosumers should divide their assets between multiple
  anonymous addresses; otherwise, each asset might be traced back to
  its prosumer based on the amount of energy that it contains.}

The prosumer can now engage in energy trading anonymously.  To find a
trade partner, it can either post an \emph{energy ask} on the bid
storage, or simply search the storage for an acceptable \emph{energy
  bid}.  To post an energy ask, the prosumer first proves to the
storage service---without revealing its original identity---that it
owns a production asset stored at an anonymous address.  It can then
post the energy ask \circled{4}, which contains an anonymous
communication identifier\footnote{We discuss communication anonymity later.}, a price, and a reference
to the production asset.  If another prosumer, who is seeking to buy energy,
finds the ask acceptable it can contact the seller using the
communication identifier included in the ask.

The seller and buyer can execute the trade by creating an energy and
financial transaction together \circled{5}, and recording it on the
ledger.  This transaction transfers the production asset from the
seller to the buyer, and a \emph{financial asset} and an \emph{energy
  consumption asset} from the buyer to the seller.  A financial asset
represents a certain amount of money, while a consumption asset
represents a permission to buy a certain amount of energy, which is
used to enforce safety requirements similarly to production assets.

Finally, the selling prosumer deposits the financial and consumption
assets to its smart meter using an energy and financial transaction.
To ensure that the prosumer remains anonymous, it transfers the assets
to an anonymous address that is randomly generated and controlled by
the smart meter \circled{6}.  Once the smart meter has received the
assets, it credits the financial amount to and deducts the energy
amount from the prosumer for billing purposes.  To enforce safety
requirements, the prosumer is required to always deposit the same amount of
consumption assets as the amount of production assets withdrawn at the
beginning; otherwise, unaccounted assets could be used to trade
excessive amounts.

\Doug{Abhishek, it seems like we might need a transition here since the previous paragraph seemed to wrap up the ``Overview of the Trading Process'' subsection above.  It's not clear how the rest of this text ties in with that discussion.  Perhaps we could use subsubsections to make these parts more distinct?}
\Aron{It did not wrap up. We have only discussed the process from the seller's perspective.}
Second, consider a prosumer who would like to buy energy from another prosumer.
Since the trading process is very similar to the case of the selling prosumer, we will discuss only the differences.
In the first step, the prosumer tries to withdraw a financial asset and an energy consumption asset from its smart meter.
If the prosumer has the consumption capacity and good financial standing, the smart meter transfers the assets to the prosumer and adds the financial amount to the prosumer's bill.

After transferring the assets through a mixing service, the prosumer is ready to post an energy bid on the bid service.
To do so, it first proves the ownership of both the financial asset and the consumption asset to the service, and then posts the energy bid, which includes an anonymous communication identifier.
If a partner is found, the trade is executed as described above, with the prosumer playing the role of the buyer this time.

Finally, the prosumer deposits the purchased energy production asset to the anonymous address of its smart meter,
which credits the energy amount to the prosumer, for billing purposes.
Note that if the prosumer has not spent all of its financial assets, then the remainder may also be deposited back to the smart meter.

\input{transactions.tex}

\input{services.tex}