### 2017-06-17 papers/IoT2017/concl.tex

\section{Conclusion}
\label{sec:concl}

\Doug{Abhishek, as always, it would be best if we could revise this summary of the paper's contents to be a summary of lessons learned based upon our experiences to date!}
As the complexity of power systems increases due to the evolution of
power grids, decentralized transactive-energy IoT systems are emerging
to tackle this complexity. Ironically, these decentralized systems
also give rise to new privacy challenges,
\Doug{Abhishek, please briefly list the key privacy challenges we cover in this paper.}
which are exacerbated by the stringent safety and
security requirements of power systems.

This paper describes \change{how we address these challenges via}{\emph{Privacy-preserving Energy Transactions} (PETra),} an
innovative solution for anonymous energy trading within a transactive
microgrid.  PETra builds on distributed ledgers, such as
blockchains, and proven techniques for anonymity, such as mixing
services and onion routing.  We described the workflow of
anonymous energy trading and explained the novel transactions and
services used in PETra.  Finally, we discussed how PETra
 satisfies security, safety, and privacy requirements.  In
future work, we will provide rigorous proofs of satisfying these
requirements, and we will present an implementation of PETra with
performance results.