### 2017-06-20 papers/IoT2017/prelim.tex

%!TEX root = paper.tex
\section{System Model and Requirements}

This section describes a basic system model of transactive IoT
microgrids and formulates security, safety, and privacy requirements.
A microgrid is a collection of prosumers (residential nodes) that are
arranged within the same distribution feeder and support exchange of
power between them. A prosumer node typically includes a smart
inverter and a smart meter, which control the flow of power into and
out of the prosumer.

A microgrid also typically contains a set of protection nodes that are
responsible for isolating faults on the feeder.  The
\emph{Distribution System Operator} (DSO) operations a set of
switching nodes to control the connection of the microgrid to the rest
of the distribution system. The DSO is responsible for regulating the
net electric power into and out of the microgrid. Starting from this
model, we next introduce the transactive microgrid model.

\subsection{Transactive Microgrid System Model}
\Abhishek{This is the perfect paragraph to define transactive
  microgrid as a special application of IoT spread over a large
  area. Perhaps we can find a citation that makes this point.}  We
describe a basic system model of decentralized transactive IoT
microgrids.  We discuss the following components: a distributed ledger
for recording transactions, a bid storage service that facilitates
finding trade partners, a microgrid controller for regulating the
microgrid load, and smart meters for measuring the prosumers' energy
production and consumption.

Figure~\ref{fig:softwareArchitecture} shows the a decentralized
transactive microgrid with PETra.  In this figure, components marked
red are introduced to provide privacy in a safe and secure manner,
whereas components marked in blue are typical elements of a
decentralized transactive microgrid.
\begin{figure}[h!]
\center
\begin{tikzpicture}[x=8cm, y=0.7cm, font=\small,
  nodeStyle/.style={rounded corners=0.1cm, drop shadow={shadow xshift=0.05cm, shadow yshift=-0.05cm, fill=black}}
]
\draw [nodeStyle, fill=red!10]  (0, 1.65) rectangle    (1, 2.25) node [midway, align=center] {Communication anonymity (e.g., onion routing)};

\draw [nodeStyle, fill=blue!10] (0, 2.4) rectangle    (1, 3.0) node [midway, align=center] {Distributed ledger (e.g., blockchain)};

\draw [nodeStyle, fill=red!10]  (0, 3.15) rectangle (0.45, 4.05) node [midway, align=center] {Transaction anonymity\\[-0.2em](mixing service)};
\draw [nodeStyle, left color=blue!10, right color=red!10] (0.47, 3.15) rectangle (0.72, 4.05) node [midway, align=center] {Bid storage};
\draw [nodeStyle, left color=blue!10, right color=red!10] (0.74, 3.15) rectangle (1, 4.05) node [midway, align=center] {Active\\[-0.2em]smart meter};

\draw [nodeStyle, fill=red!10] (0, 4.2) rectangle (0.49, 5.1) node [midway, align=center] {Anonymous trading\\[-0.2em]workflow for prosumers};
\draw [nodeStyle, fill=blue!10] (0.51, 4.2) rectangle (1, 5.1) node [midway, align=center] {Microgrid controller};
\end{tikzpicture}
\caption{The Architecture of a Decentralized Transactive Microgrid with PETra.}
\label{fig:softwareArchitecture}
\end{figure}

\subsubsection{Distributed Ledger}
This ledger permanently stores transactions that specify energy
trades, change regulatory policies for the microgrid, etc.  For
providing security and safety, it is crucial that transactions be
immutable, \emph{i.e.}, after a transaction has been recorded, it
cannot be modified or removed from the ledger.  To enhance fault
tolerance, however, the ledger should also be distributed.

Since a distributed ledger is maintained by multiple nodes, a key
requirement is reaching consensus on which transactions are valid and
stored on the ledger.  Moreover, this consensus must be reached
quickly and reliably, even in the presence of faulty or malicious
(\emph{e.g.}, compromised) ledger nodes.  This paper assumes that a
distributed ledger service is available, but makes no assumptions
about the ledger implementation, such as the particulars of the
consensus algorithm.  In practice, a distributed ledger can be
implemented using, \emph{e.g.}, \emph{blockchains} with proof-of-stake
consensus or a practical Byzantine fault tolerance
algorithm~\cite{castro1999practical}.

\subsubsection{Bid Storage Service}
Although prosumers trade energy with each other directly (\emph{i.e.},
without a middleman), for the sake of scalability, we need a service
that enables prosumers to find trade partners.  We assume that there
is a bid storage service that allows prosumers to post and read energy
\emph{bids} and \emph{asks}.\footnote{A \emph{bid} is an offer to buy
  at a certain price, while an \emph{ask} is an offer to sell at a
  certain price.}  This service relieves prosumers from contacting a
large number of potential trade partners since they only communicate
with the service to discover trade partners.  To enhance scalability
and reliability, this service can also be implemented in a distributed
manner, using multiple nodes.

\subsubsection{Microgrid Controller (Distribution System Operator)}
We assume the existence of a controller at the DSO level that
regulates the total load that the microgrid should present to the
distribution system.  The controller first predicts load in the
microgrid based on (1) bids and asks in the bid storage and (2)
outstanding energy trades in the ledger.  By combining this
information with the prediction for the rest of the grid, the
controller produces a control signal that specifies how much the
microgrid load should be decreased or increased.  Based on this
signal, the controller then updates the price policy for the microgrid
to influence energy production and consumption.  We also assume the
presence of a secondary controller that balances voltage and frequency
in the microgrid.

\subsubsection{Smart Meters}
To measure the prosumers' energy production and consumption, a smart
meter must be deployed at each prosumer.  In practice, these smart
meters must be tamper resistant to prevent prosumers from ``stealing
electricity'' by tampering with their meters.  After a smart meter has
measured the net amount of energy consumed by the prosumer in some
time interval, it can send this information to the DSO for billing
purposes.

\subsection{Requirements}
We now discuss the security, safety, and privacy requirements that
must be satisfied by a transactive energy IoT system.

\subsubsection{Security}
Security requirements ensure primarily that prosumers are billed
correctly, but they also provide necessary prerequisite properties for
safety.
More specifically, they require that
\begin{itemize}[noitemsep,topsep=-\parskip]
\item prosumers are billed correctly based on the energy prices set by
  the DSO, their energy trades, and their actual energy production and
  consumption measured by the smart meters,
\item prosumers or outside attackers cannot change microgrid
  regulatory policies that are set by the DSO, 
\item prosumers cannot back out of trades unilaterally, and they
  cannot tamper with other prosumers' trading or bidding,
\item financial and physical impact of compromised or faulty nodes is
  limited, and nodes can be banned by the DSO. 
\end{itemize}

\subsubsection{Safety}
A careless or malicious prosumer may destabilize the grid by promising
to produce (or consume) a large amount of energy, but failing to actually
produce (or consume) it.  A significant difference between promised and
actual energy production (or consumption) can result in a large gap
between the aggregate production and consumption of the microgrid.
A large gap threatens the stability of not only the microgrid but also the main
power grid.  Therefore, prosumers should not be able to trade large
amounts of energy that they are unlikely to deliver.
Specifically, we require that 
\begin{itemize}[noitemsep,topsep=-\parskip]
\item the net amount of energy sold (or bought) by a prosumer is upper
  bounded (by a limit set by the DSO), where the net amount of
  energy sold is the difference between the amount of energy sold and
  bought by the prosumer, and the net amount of energy bought is
  defined analogously. %,
\Aron{If we have space, add back the safety requirement on the bid storage (and also add back the relevant parts of the description of the trading workflow and the ``analysis.''}
%\item the energy bids and asks posted by a prosumer are limited in a similar way.
\end{itemize}
In practice, the DSO can set the limits based on the prosumers' production and consumption capacities.

\subsubsection{Privacy} 
Privacy requirements ensure that the prosumers' privacy is not
compromised when they participate in energy trading.  We use
non-transactive smart metering as a baseline, and we require that
the transactive system does not leak any additional information
compared to this baseline.  More specifically, we require that
\begin{itemize}[noitemsep,topsep=-\parskip]
\item only the corresponding smart meter and the DSO may gain
  information regarding the amount of energy produced, consumed,
  bought, or sold by a prosumer,\footnote{Although this requirement is
    impossible to satisfy if all other prosumers may collude against
    one target, we can assume that the majority of prosumers are
    non-colluding.}
\item only the prosumer may know which bids and asks it has posted,
  and no one can know who traded energy with whom.
\end{itemize}