### 2017-06-20 papers/IoT2017/solution.tex

%!TEX root = paper.tex
\section{Privacy-preserving Energy Transactions}

This section describes \emph{Privacy-preserving Energy Transactions}
(PETra), which is our solution for providing privacy to prosumers in a
transactive energy IoT system, without compromising grid safety and
security.
\begin{figure*}[h]
\centering
\begin{tikzpicture}[x=3.8cm, y=1.6cm, font=\small,
  system/.style={draw, align=center, rounded corners=0.1cm, fill=black!5},
  entity/.style={draw, align=center, rounded corners=0.1cm},
  asset/.style={midway, align=center},
  transfer/.style={->, >=stealth, shorten <=0.05cm, shorten >=0.05cm},
]
\node[entity] (smartmeter) at (0, 1) {smart meter};
\node[entity] (prosumer1) at (1, 1) {prosumer};
\node[system] (mixing1) at (1.9, 1) {mixing\\service};
\node[entity] (prosumer2) at (3, 1) {prosumer\\(anonymous address)};
\node[system] (bidstorage) at (4, 0.5) {bid\\storage};
\node[entity] (partner) at (3, 0) {other prosumer};
\node[entity] (prosumer3) at (1.5, 0) {prosumer\\(anonymous address)};
\node[entity] (smartmeter2) at (0, 0) {smart meter\\(anonymous address)};

\draw[transfer] (smartmeter) -- node [asset, above] {\circled{1} energy\\production asset} (prosumer1);
\draw[transfer] (prosumer1) -- node [asset, above] {\circled{2} energy\\production asset} (mixing1);
\draw[transfer] (mixing1) -- node [asset, above] {\circled{3} energy\\production asset} (prosumer2);
\draw[transfer, dashed] (prosumer2) -- node [asset, above right, xshift=-1.5em, yshift=0.5em] {\circled{4} energy production asset,\\energy ask} (bidstorage);
\draw[transfer, dashed] (bidstorage) -- node [asset, below right] {energy\\ask} (partner);
\draw[transfer, bend right=30] (partner) to node [asset, below] {\circled{5} financial asset,\\energy consumption asset} (prosumer3);
\draw[transfer, bend right=50] (prosumer2) to node [asset, right, yshift=0.25em] {\circled{5} energy\\production asset} (partner);
\draw[transfer] (prosumer3) -- node [asset, above] {\circled{6} financial asset,\\
energy consumption asset} (smartmeter2);
\end{tikzpicture}
\caption{Simplified Overview of the Flow of Assets from the Perspective of a Prosumer Who Sells Energy.}
\label{fig:sellFlow}
\end{figure*}

\subsection{Overview of the Trading Workflow}
We now provide a semi-formal description of the energy trading
workflow from the prosumers' perspective.  Subsequent subsections
describe the assets, transactions, and services used for trading in
more detail.

\subsubsection{Energy Selling Workflow}
Consider a prosumer who wishes to sell energy to another prosumer, as
shown in Figure~\ref{fig:sellFlow}\footnote{To prevent
  de-anonymization, a prosumer should use multiple addresses and
  multiple rounds of mixing, which we have omitted from the figure for
  clarity of presentation.}  \Abhishek{We need to mention what is the
  expected deadline by which this sequence of task should finish.}  As
its first step, the prosumer withdraws an \emph{energy production
  asset} from its smart meter.  An energy production asset represents
a permission to sell a certain amount of energy, and it is used to
enforce safety requirements.  If the prosumer has sufficient unsold
production capacity, the smart meter creates and transfers a
production asset to the prosumer using a \emph{smart meter
  transaction} \circled{1}, which is recorded on the distributed
ledger.

At this point, the production asset can still be traced back to the
prosumer since the ledger is public.  To achieve anonymity, the
prosumer uses a \emph{mixing service}, which could be implemented as a
decentralized protocol, such as
CoinShuffle~\cite{ruffing2014coinshuffle} or
Xim~\cite{bissias2014sybil}.  The prosumer transfers the production
asset to the mixing service using an \emph{energy and financial
  transaction} \circled{2}, which is also recorded on the distributed
ledger.  In turn, the mixing service transfers the production asset to
an \emph{anonymous address} \circled{3}, which is randomly generated
and controlled by the prosumer.\footnote{The concept of \emph{address}
  varies between distributed ledgers, but PETra could be implemented
  using any popular blockchain, such as Bitcoin and Ethereum.
  Specifically, we use the term address to denote a possible
  destination for asset transfers.  Assets that have been transferred
  to an address can be used only by someone who ``controls'' the
  address (typically, the one who generated it), which usually means
  knowing a private key that corresponds to the address.}  Since the
mixing service transfers assets from multiple prosumers to multiple
anonymous addresses at the same time, and the anonymous addresses were
generated at random by the prosumers, the assets cannot be traced back
to the original prosumers after mixing.\footnote{Note that prosumers
  should divide their assets between multiple anonymous addresses;
  otherwise, each asset might be traced back to its prosumer based on
  the amount of energy that it contains.}

The prosumer can now engage in energy trading anonymously.  To find a
trade partner, it can either post an \emph{energy ask} on the bid
storage, or simply search the storage for an acceptable \emph{energy
  bid}.  To post an energy ask, the prosumer first proves to the
storage service---without revealing its original identity---that it
owns a production asset stored at an anonymous address.  Proving
ownership prevents the prosumer from ``spamming'' the storage service
with bogus asks.  The prosumer can then post the energy ask
\circled{4}, which contains an anonymous communication
identifier\footnote{We discuss communication anonymity later.}, a
price, and a reference to the production asset.  If another prosumer,
who is seeking to buy energy, finds the ask acceptable it can contact
the seller using the communication identifier included in the ask.

The seller and buyer can execute the trade by creating an energy and
financial transaction together \circled{5}, and recording it on the
ledger.  This transaction transfers the production asset from the
seller to the buyer, and a \emph{financial asset} and an \emph{energy
  consumption asset} from the buyer to the seller.  A financial asset
represents a certain amount of money, while a consumption asset
represents a permission to buy a certain amount of energy, which is
used to enforce safety requirements similarly to production assets.

Finally, the selling prosumer deposits the financial and consumption
assets to its smart meter using an energy and financial transaction.
To ensure that the prosumer remains anonymous, it transfers the assets
to an anonymous address that is randomly generated and controlled by
the smart meter \circled{6}.  Once the smart meter has received the
assets, it credits the financial amount to and deducts the energy
amount from the prosumer for billing purposes.  To enforce safety
requirements, the prosumer is required to always deposit the same
amount of consumption assets as the amount of production assets
withdrawn at the beginning; otherwise, unaccounted assets could be
used to trade excessive amounts.

\subsubsection{Energy Buying Workflow}
Consider a prosumer who would like to buy energy from another
prosumer.  Since the trading workflow is very similar to the case of
the selling prosumer, we will discuss only the differences.  In the
first step, the prosumer tries to withdraw a financial asset and an
energy consumption asset from its smart meter.  If the prosumer has
the consumption capacity and good financial standing, the smart meter
transfers the assets to the prosumer and adds the financial amount to
the prosumer's bill.

After transferring the assets through a mixing service, the prosumer
is ready to post an energy bid on the bid service.  To do so, it first
proves the ownership of both the financial asset and the consumption
asset to the service, and then posts the energy bid, which includes an
anonymous communication identifier.  If a partner is found, the trade
is executed as described above, with the prosumer playing the role of
the buyer this time.

Finally, the prosumer deposits the purchased energy production asset
to the anonymous address of its smart meter, which credits the energy
amount to the prosumer, for billing purposes.  Note that if the
prosumer has not spent all of its financial assets, then the remainder
may also be deposited back to the smart meter.

\input{transactions.tex}

\input{services.tex}