### 2017-06-21 papers/IoT2017/concl.tex
update

%!TEX root = paper.tex
\section{Concluding Remarks}
\label{sec:concl}

\Doug{Abhishek, as always, it would be best if we could revise this summary of the paper's contents to be a summary of lessons learned based upon our experiences to date!}
As the complexity of power systems increases due to the evolution of
power grids, decentralized transactive-energy IoT systems are emerging
to tackle this complexity. Ironically, these decentralized systems
also give rise to new privacy challenges, such as leakage of energy usage patterns, including the possibility of inference of future state of a prosumer, 
\Doug{Abhishek, please briefly list the key privacy challenges we cover in this paper.}
which are exacerbated by the stringent safety and
security requirements of power systems.

This paper describes how we address these challenges via
\emph{Privacy-preserving Energy Transactions} (PETra), which is our
innovative solution for anonymous energy trading within a transactive
microgrid.  PETra builds on distributed ledgers, such as blockchains,
and proven techniques for anonymity, such as mixing services and onion
routing.  We described the workflow of anonymous energy trading and
explained the novel transactions and services used in PETra.  Finally,
we discussed how PETra satisfies security, safety, and privacy
requirements.  In future work, we will provide rigorous proofs of
satisfying these requirements.
% and we will present an implementation
%of PETra with performance results.