### 2017-06-21 papers/IoT2017/concl.tex
final

%!TEX root = paper.tex
\section{Concluding Remarks}
\label{sec:concl}

\Doug{Abhishek, as always, it would be best if we could revise this summary of the paper's contents to be a summary of lessons learned based upon our experiences to date!}
As the complexity of power systems increases due to the evolution of
power grids, decentralized transactive-energy IoT systems are emerging
to tackle this complexity. Ironically, these decentralized systems
also give rise to new privacy challenges, such as the potential leakage of energy usage patterns, including the possibility of inferring the future state of a prosumer.
These challenges are exacerbated by the stringent safety and
security requirements of power systems.

This paper describes 
\emph{Privacy-preserving Energy Transactions} (PETra), our
innovative solution for anonymous energy trading within a transactive
microgrid.  PETra builds on distributed ledgers, such as blockchains,
and proven techniques for anonymity, such as mixing services and onion
routing.  We described the workflow of anonymous energy trading and
explained the novel transactions and services used in PETra.  Finally,
we discussed how PETra satisfies security, safety, and privacy
requirements.  In future work, we will provide rigorous proofs of
satisfying these requirements.
% and we will present an implementation
%of PETra with performance results.