### 2017-07-21 delieverables/July2017/test-network-setup-seq-diag.tex

\documentclass{standalone}

\usepackage{pgf-umlsd}
\usepackage{adjustbox}

\begin{document}
	\trimbox{0cm -.5cm -0.75cm -0.5cm}{

	\begin {sequencediagram}
%	\newinst{name}{:Instance}
	\newthread {a}{: Manager Install}
%	\newinst {b }{: Manager}
	\newinst [1]{d}{: Bootnodes }
%	\newinst [1]{ch}{: Client Hosts }
	\newinst [2]{c}{: Clients }
%	\newinst [3]{mh}{: Miner Hosts }
	\newinst [3]{m}{: Miners }

%	\newinst [5]{dh}{: Discovery Hosts }
\begin{sdblock}{Creation}{Creating all Ethereum related processes}
	\begin{call}{a}{create node(s)}{d}{KEY:IP:PORT (s)}
	\end{call}

	\begin{call}{a}{create miner(s)}{m}{KEY:IP:PORT (s)}
	\end{call}
	\begin{call}{a}{create client(s)}{c}{KEY:IP:PORT (s)}
	\end{call}
	\begin{call}{a}{create Genesis Block, with default credits}{a}{Genesis Block}
	\end{call}

	\begin{call}{a}{copy Genesis Block \& bootnode info}{m}{}
	\end{call}
	\begin{call}{a}{copy Genesis Block \& bootnode info}{c}{}
	\end{call}
\end{sdblock}

\begin{sdblock}{Starting}{Starting all Ethereum related processes}
	\begin{call}{a}{start node(s)}{d}{}
	\end{call}
	\begin{call}{a}{start Miner(s)}{m}{}
	\end{call}
	\begin{call}{a}{Start Client(s)}{c}{}
	\end{call}
\end{sdblock}

\begin{sdblock}{Stopping}{Stopping all Ethereum related processes}
	\begin{call}{a}{stop Client(s)}{c}{}
	\end{call}
	\begin{call}{a}{stop Miner(s)}{m}{}
	\end{call}
	\begin{call}{a}{stop node(s)}{d}{}
	\end{call}
\end{sdblock}

\begin{sdblock}{Removal}{Removing all Ethereum related processes}
	\begin{call}{a}{remove Client(s)}{c}{}
	\end{call}
	\begin{call}{a}{remove Miner(s)}{m}{}
	\end{call}
	\begin{call}{a}{remove node(s)}{d}{}
	\end{call}
\end{sdblock}

	\end {sequencediagram}
}
\end{document}