### 2017-08-22 docs/July2017/clients-commands.tex

\subsection{Clients}

Clients are the nodes in the test network that have the potential to interact with the Blockchain that are NOT miners. Note: Mining is a separate process that we dedicate nodes to, so that IoT devices (Smart Meters, etc.) do not need to run intensive computation calculations.

\subsubsection{Create}
The following command allows you to create the clients specified in <network-config.json> on the test network. This command evaluates the number of hosts given, the number of client types specified, the number of clients per type, and the starting port for all clients.
\begin{lstlisting}[language=bash]
$ ./manager clients create file=network-config.json
\end{lstlisting}

\subsubsection{Distribute}
The following command allows you to distribute the <static-nodes.json> that contains the bootnode(s) connection information to all Clients. Note: This includes all types declared within the 'clients' section of the <network-config.json> file, such as: Prosumer, Miners, etc.
\begin{lstlisting}[language=bash]
$ ./manager clients create file=network-config.json nodes=static-nodes.json
\end{lstlisting}

\subsubsection{Start}
The following command allows you to start all the clients on the test network declared in the provided <network-config.json> file.
\begin{lstlisting}[language=bash]
$ ./manager clients create file=network-config.json
\end{lstlisting}

\subsubsection{Stop}
The following command allows you to stop all the clients on the test network declared in the provided <network-config.json> file.
\begin{lstlisting}[language=bash]
$ ./manager clients create file=network-config.json
\end{lstlisting}

\subsubsection{Delete}
The following command allows you to delete all the clients' data on the test network declared in the provided <network-config.json> file.
\begin{lstlisting}[language=bash]
$ ./manager clients create file=network-config.json
\end{lstlisting}