### 2017-09-27 docs/papers/m4iot-2017/paper.tex

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage[mode=buildnew]{standalone}% requires -shell-escape
\usepackage{xcolor}

\usepackage{hyperref}

\usepackage{tikz}
\usetikzlibrary{calc}

\usepackage{standalone}

\usepackage{listings}

\usetikzlibrary{shadows}
\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
%\pgfplotsset{compat=1.10}

\usepackage{multirow}

\lstset{
	basicstyle=\ttfamily,
	frame=single
}

% Put % before of what you want disabled

% Select what to do with todonotes: 
\usepackage[textsize=tiny]{todonotes} % notes not showed
%\usepackage[draft, textsize=scriptsize, textwidth=1.55cm]{todonotes}   % notes showed
\newcommand{\Aron}[1]{\todo{\textbf{Aron}: #1}}
\newcommand{\Abhishek}[1]{\todo{\textbf{Abhishek}: #1}}
\newcommand{\Mike}[1]{\todo[backgroundcolor=blue!25]{\textbf{Mike}: #1}}
\newcommand{\Doug}[1]{\todo{\textbf{Doug}: #1}}

\newcommand{\AronOld}[1]{\todo[backgroundcolor=yellow!15]{\textbf{Aron\\Addressed}: #1}}
\newcommand{\MikeOld}[1]{\todo[backgroundcolor=yellow!15]{\textbf{Mike\\Addressed}: #1}}
\newcommand{\DougOld}[1]{\todo[backgroundcolor=yellow!15]{\textbf{Doug\\Addressed}: #1}}
%\newcommand{\AbhishekOld}[1]{\todo[backgroundcolor=yellow!25]{\textbf{Abhishek\\Addressed}: #1}}
\newcommand{\AbhishekOld}[1]{}
% Select what to do with command \comment:  
% \renewcommand{\comment}[1]{}  %comment not showed
 \renewcommand{\comment}[1] {\par {\bfseries \color{blue} #1 \par}} %comment showed

% Copyright
%\setcopyright{none} %\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
\acmConference[M4IoT]{4th Workshop on Middleware and Applications for the Internet of Things}{December 2017}{Las Vegas, USA} 
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\begin{document}
\setlength{\marginparwidth}{1.5cm}

 \title[PlaTIBART: a Platform for Transactive IoT Blockchain Applications with Repeatable Testing]{PlaTIBART: a Platform for Transactive IoT Blockchain Applications with Repeatable Testing}

%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
 % \texttt{acmart.pdf} document}

\author{Michael A. Walker, Abhishek Dubey, Aron Laszka, and Douglas C. Schmidt}
\affiliation{\institution{Vanderbilt University}}
\email{{michael.a.walker.1, abhishek.dubey, a.laszka, d.schmidt}@vanderbilt.edu, }

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{M. Walker et al.}

\begin{abstract}
{\em
With the advent of blockchain-enabled IoT applications, there is an increased need for related software patterns, middleware concepts, and testing practices to ensure adequate quality and productivity. IoT and blockchain each provide different design goals, concepts, and practices that must be integrated, including the distributed actor model and fault tolerance from IoT and transactive information integrity over untrustworthy sources from blockchain.
\AbhishekOld{for example, .. are we referring to distributed actor model and transactive information integrity over untrustworthy sources. It will be good to be precise}
Both IoT and blockchain are emerging technologies and both lack codified patterns and practices for development of applications when combined.
\AbhishekOld{This is first time we have used the world cyber-physical here}
This paper describes PlaTIBART, which is a platform for transactive IoT blockchain applications with repeatable testing that combines the Actor pattern (which is a commonly used model of computation in IoT\AbhishekOld{How does this relate to IoT, you can say a commonly used model of computation in IoT}) together with a custom Domain Specific Language (DSL) and test network management tools. We show how PlaTIBART has been applied to develop, test, and analyze fault-tolerant \AbhishekOld{Fault-tolerant is being described very late. you should describe this earlier as an example of the well-defined patterns} IoT blockchain applications.}

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
  <concept>
<concept_id>10011007.10011006.10011073</concept_id>
<concept_desc>Software and its engineering~Software maintenance tools</concept_desc>
<concept_significance>500</concept_significance>
</concept>
 <concept>
<concept_id>10011007.10011074.10011081.10011082.10011088</concept_id>
<concept_desc>Software and its engineering~Design patterns</concept_desc>
<concept_significance>500</concept_significance>
</concept>
  <concept>
<concept_id>10010520.10010575</concept_id>
<concept_desc>Computer systems organization~Dependable and fault-tolerant systems and networks</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[500]{Software and its engineering~Software maintenance tools}
\ccsdesc[500]{Software and its engineering~Design patterns}
\ccsdesc[300]{Computer systems organization~Dependable and fault-tolerant systems and networks}

\keywords{Internet of Things, Blockchain, Design Patterns, Testing}

%\acmBadgeR{artifacts_available}

\maketitle

\input{main-body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}