### 2017-10-16 docs/papers/iotdi2018/concl.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper, we described the design and implementation of a  transaction management platform (TMP) for IoT-based transactive microgrids. Our solution enables prosumers to trade energy without threatening their privacy or the safety of the system. Our hybrid solver approach, which combines a smart-contract based validator with an external optimizer, enables the platform to clear offers securely and efficiently. Further, the ability to trade across multiple time intervals enables participants to take full advantage of batteries, thereby smoothening the load on the main grid. Finally, the use of blockchains provides decentralized trust and consensus capabilities, which protect from  malicious actors.

\AronC{future work:
- recovery from failures: retroactively change matching in case of a physical failure

- fault-tolerant matching: safety constraints will be satisfied even if a certain number of physical elements fail (or prosumers get disconnected)

- propose alternative market mechanisms, including those that involve double auctions and its variants.}