### 2018-05-18 docs/papers/IEEEBlockchain2018/main.tex

\documentclass[conference, compsoc]{IEEEtran}

\newif\ifExtended
\Extendedfalse

\usepackage{booktabs} % For formal tables
\usepackage{tikz}
\usetikzlibrary{shadows}
\usepackage{pgfplots}
\pgfplotsset{compat=1.10}

\usepackage{colortbl}
\definecolor{Gray}{gray}{0.925}

\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[inline]{enumitem}
\usepackage{comment}
\usepackage{soul}
\usepackage{xspace}

\usepackage{amssymb}
\usepackage{amsmath}

\newcommand{\vect}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\va}{\vect{a}}
\newcommand{\vo}{\vect{o}}

\newcommand{\calA}[0]{\ensuremath{\mathcal{A}}}
\newcommand{\calC}[0]{\ensuremath{\mathcal{C}}}
\newcommand{\calO}[0]{\ensuremath{\mathcal{O}}}
\newcommand{\calP}[0]{\ensuremath{\mathcal{P}}}
\newcommand{\calR}[0]{\ensuremath{\mathcal{R}}}

\newcommand{\Natural}[0]{\ensuremath{\mathbb{N}}}
\newcommand{\Real}[0]{\ensuremath{\mathbb{R}}}

\usepackage{amsthm}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

%-----Solidity Environment--------------------------------
\usepackage{listings, xcolor}
\definecolor{verylightgray}{rgb}{.97,.97,.97}
\lstdefinelanguage{Solidity}{
	keywords=[1]{anonymous, assembly, assert, balance, break, call, callcode, case, catch, class, constant, continue, contract, debugger, default, delegatecall, delete, do, else, event, export, external, false, finally, for, function, gas, if, implements, import, in, indexed, instanceof, interface, internal, is, length, library, log0, log1, log2, log3, log4, memory, modifier, new, payable, pragma, private, protected, public, pure, push, require, return, returns, revert, selfdestruct, send, storage, struct, suicide, super, switch, then, this, throw, transfer, true, try, typeof, using, value, view, while, with, addmod, ecrecover, keccak256, mulmod, ripemd160, sha256, sha3}, % generic keywords including crypto operations
	keywordstyle=[1]\color{blue}\bfseries,
	keywords=[2]{address,..., bool, byte, bytes, bytes1, bytes2, bytes3, bytes4, bytes5, bytes6, bytes7, bytes8, bytes9, bytes10, bytes11, bytes12, bytes13, bytes14, bytes15, bytes16, bytes17, bytes18, bytes19, bytes20, bytes21, bytes22, bytes23, bytes24, bytes25, bytes26, bytes27, bytes28, bytes29, bytes30, bytes31, bytes32, enum, int, int8, int16, int24, int32, int40, int48, int56, int64, int72, int80, int88, int96, int104, int112, int120, int128, int136, int144, int152, int160, int168, int176, int184, int192, int200, int208, int216, int224, int232, int240, int248, int256, mapping, string, uint, uint8, uint16, uint24, uint32, uint40, uint48, uint56, uint64, uint72, uint80, uint88, uint96, uint104, uint112, uint120, uint128, uint136, uint144, uint152, uint160, uint168, uint176, uint184, uint192, uint200, uint208, uint216, uint224, uint232, uint240, uint248, uint256, var, void, ether, finney, szabo, wei, days, hours, minutes, seconds, weeks, years},	% types; money and time units
	keywordstyle=[2]\color{teal}\bfseries,
	keywords=[3]{block, blockhash, coinbase, difficulty, gaslimit, number, timestamp, msg, data, gas, sender, sig, value, now, tx, gasprice, origin},	% environment variables
	keywordstyle=[3]\color{violet}\bfseries,
	identifierstyle=\color{black},
	sensitive=false,
	comment=[l]{//},
	morecomment=[s]{/*}{*/},
	commentstyle=\color{gray}\ttfamily,
	stringstyle=\color{red}\ttfamily,
	morestring=[b]',
	morestring=[b]"
}
\lstset{
	language=Solidity,
	backgroundcolor=\color{verylightgray},
	extendedchars=true,
	basicstyle=\footnotesize\ttfamily,
	showstringspaces=false,
	showspaces=false,
	numbers=left,
	numberstyle=\footnotesize,
	numbersep=1pt,
	tabsize=2,
	breaklines=true,
	showtabs=false,
	captionpos=b,
	xleftmargin=.5em,
	xrightmargin=.5em
}

\newcommand{\change}[2]{\st{#1} {\textcolor{blue}{#2}}}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[textsize=tiny, textwidth=1.5cm,disable]{todonotes}
\newcommand{\TODO}[1]{\todo[inline]{#1}}
\newcommand{\ad}[1]{\todo[color=yellow!50, linecolor=black!50]{\textbf{Abhishek}: #1}}
\newcommand{\Aron}[1]{\todo[color=blue!25, linecolor=black!50]{\textbf{Aron}: #1}}
\newcommand{\Anastasia}[1]{\todo[color=pink, linecolor=black!50]{\textbf{Anastasia}: #1}}
\newcommand{\Scott}[1]{\todo[color=orange!30, linecolor=black!50]{\textbf{Scott}: #1}}

%\usepackage{hyperref}

\usepackage{url}
\graphicspath{{./figs/}}

\usepackage{titlesec}

\titlespacing\section{0pt}{8pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{8pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}
\titlespacing\subsubsection{0pt}{8pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\Platform}{\textit{SolidWorx}\xspace}

\begin{document}

\setlength{\marginparwidth}{1.4cm}

\thispagestyle{plain}
\pagestyle{plain}

\title{\Platform: A Resilient and Trustworthy Transactive Platform for\\Smart and Connected Communities}
\author{
\IEEEauthorblockN{Scott Eisele}
\IEEEauthorblockA{Vanderbilt University \\
Nashville, TN, USA \\ 
scott.r.eisele@vanderbilt.edu
}

\and
\IEEEauthorblockN{Aron Laszka}
\IEEEauthorblockA{University of Houston\\
Houston, TX, USA \\
alaszka@houston.edu}
\and
\IEEEauthorblockN{Anastasia Mavridou}
\IEEEauthorblockA{Vanderbilt University\\
Nashville, TN, USA \\
anastasia.mavridou@vanderbilt.edu}
\and
\IEEEauthorblockN{Abhishek Dubey}
\IEEEauthorblockA{Vanderbilt University \\
Nashville, TN, USA \\ 
abhishek.dubey@vanderbilt.edu
}
}

\maketitle

\input{summary}
\setlength{\belowcaptionskip}{-10pt}
\section{Introduction}
\label{sec:intro}
\input{introduction}

\section{Problem Formulation}
\label{sec:problem}
\input{problem}

\section{\Platform: A Decentralized Transaction Management Platform}
\label{sec:solution}
\input{solution}
\subsection{Smart Contract}
\input{fsolidm}

%\subsection{Workflow}
%\label{sec:analysis}
%\input{workflow}

\subsection{Analysis}
\label{sec:analysis}
\input{analysis}
\section{Case Studies}
\label{sec:results}
\input{results}
\section{Related Research}
\label{sec:related}
\input{related}
\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

\vspace{0.5em}
\noindent\textbf{Acknowledgement:}
This work was funded in part by a grant from Siemens, CT and in part by a grant from NSF under award number CNS-1647015.

\let\oldbibliography\thebibliography
\renewcommand{\thebibliography}[1]{\oldbibliography{#1}
\setlength{\itemsep}{0pt}} %Reducing spacing in the bibliography.

\bibliographystyle{IEEEtran}
%\setlength{\bibsep}{0.0pt}
\small{
\bibliography{references} 
}

%\newpage
%\tableofcontents

\end{document}